\documentclass{./../../Latex/homework}
\begin{document}
\thispagestyle{plain}
\myheader{Homework 6 Problems}

%%%%%%%%%%%%%% Exercise 10.5
\subsection*{Exercise 10.5}

\begin{enumerate}

% Question 1
\item[1.] Find the derivatives of:
\begin{tasks}(3)
\task $y=e^{2 t+4}$
\task $y=e^{1-9 t}$
\task $y=e^{t^{2}+ 1}$
\task $y=5 e^{2-t^{2}}$
\task $y=e^{a x^{2}+b x+c}$
\task $y=x e^{x}$
\task $y=x^{2} e^{2 x}$
\task $y=a x e^{b x+c}$	
\end{tasks}

% Question 3
\item[3.] Find the derivatives of:
\begin{tasks}(3)
\task $y=\ln \left(7 t^{5}\right)$
\task $y=\ln \left(a t^{\circ}\right)$
\task $y=\ln (t+19)$
\task $y=5 \ln (t+1)^{2}$
\task $y=\ln x-\ln (1+x)$
\task $y=\ln \left[x(1-x)^{8}\right]$
\task $y=\ln \left(\frac{2 x}{1+x}\right)$
\task $y=5 x^{4} \ln x^{2}$
\end{tasks}

% Question 7
\item[7.] Find the derivatives of the following by first taking the natural log of both sides:
\begin{enumerate}
\item $y=\dfrac{3 x}{(x+2)(x+4)}$
\item $y=\left(x^{2}+3\right) e^{x^{2}+1}$
\end{enumerate}
\end{enumerate}

%%%%%%%%%%%%%% Exercise 7.4
\subsection*{Exercise 7.4}

\begin{enumerate}

% Question 1
\item[1.] Find $\partial y / \partial x_{1}$ and $\partial y / \partial x_{2}$ for each of the following functions:
\begin{tasks}(2)
\task[(a)] $y=2 x_{1}^{3}-11 x_{1}^{2} x_{2}+3 x_{2}^{2}$
\task[(d)]$y=\dfrac{5 x_{1}+3}{x_{2}-2}$ 
\end{tasks}

% Question 2
\item[2.] Find $f_{x}$ and $f_{y}$ from the following:
\begin{tasks}(2)
\task[(a)] $f(x, y)=x^{2}+5 x y-y^{3}$
\task[(b)] $f(x, y)=\dfrac{2 x-3 y}{x+y}$
\end{tasks}


% Question 3
\item[3.] From the answers to Prob. 2 , find $f_{x}(1,2)$, the value of the partial derivative $f_{x}$ when $x=1$ and $y=2$, for each function.

% Question 5
\item[5.] If the utility function of an individual takes the form
$$U=U\left(x_{1}, x_{2}\right)=\left(x_{1}+2\right)^{2}\left(x_{2}+3\right)^{3}$$
where $U$ is total utility, and $x_{1}$ and $x_{2}$ are the quantities of two commodities consumed:
\begin{enumerate}
  \item Find the marginal-utility function of each of the two commodities.
  \item Find the value of the marginal utility of the first commodity when 3 units of each commodity are consumed.
\end{enumerate}

% Question 7
\item[7.] Write the gradients of the following functions:
\begin{enumerate}
\item $f(x, y, z)=x^{2}+y^{3}+z^{4}$
\item $f(x, y, z)=x y z$
\end{enumerate}
\end{enumerate}

%%%%%%%%%%%%%% Exercise 8.1
\subsection*{Exercise 8.1}

\begin{enumerate}

% Question 1
\item[1.] Find the differential $d y$, given:
\begin{itemize}
\item[(a)] $y=-x\left(x^{2}+3\right)$
\end{itemize}

% Question 4
\item[4.] Find the point elasticity of demand, given $Q=k / P^{n}$, where $k$ and $n$ are positive constants.
\begin{enumerate}
 \item Does the elasticity depend on the price in this case?
\item In the special case where $n=1$, what is the shape of the demand curve? What is the point elasticity of demand?
\end{enumerate}

% Question 6
\item[6.] Given $Q=100-2 P+0.02 Y$, where $Q$ is quantity demanded, $P$ is price, and $Y$ is income, and given $P=20$ and $Y=5,000$, find the
\begin{enumerate}
\item Price elasticity of demand.
\item Income elasticity of demand.
\end{enumerate}

\end{enumerate}

%%%%%%%%%%%%%% Exercise 8.2
\subsection*{Exercise 8.2}

\begin{enumerate}

% Question 3
\item[3.] Find the total differential, given
\begin{itemize}
\item[(a)]$y=\dfrac{x_{1}}{x_{1}+x_{2}}$
\end{itemize}

% Question 4
\item[4.] The supply function of a certain commodity is
$$
Q=a+b P^{2}+R^{1 / 2} \quad(a<0, b>0) \quad \text { [R: rainfall] }
$$
Find the price elasticity of supply $\varepsilon_{Q P}$, and the rainfall elasticity of supply $\varepsilon_{Q R}$. 
% Question 5
\item[5.] How do the two partial elasticities in Prob. 4 vary with $P$ and $R$? In a strictly monotonic fashion (assuming positive $P$ and $R$ )?


% Question 6
\item[6.] The foreign demand for our exports $X$ depends on the foreign income $Y_{f}$ and our price level $P: X=Y_{f}^{1 / 2}+P^{-2}$. Find the partial elasticity of foreign demand for our exports with respect to our price level.

% Question 7
\item[7.] Find the total differential for each of the following functions:
\begin{tasks}
\task[(b)] $U=7 x^{2} y^{3}$
\task[(f)] $U=(x-3 y)^{3}$
\end{tasks}

\end{enumerate}

%%%%%%%%%%%%%% Exercise 8.4
\subsection*{Exercise 8.4}

\begin{enumerate}

% Question 2
\item[2.] Find the total derivative $d z / d t$, given
\begin{enumerate}
\item $z=x^{2}-8 x y-y^{3}$, where $x=3 t$ and $y=1-t$ 
\item $z=7 u+v t$, where $u=2 t^{2}$ and $v=t+1$
\item $z=f(x, y, t)$, where $x=a+b t$ and $y=c+k t$
\end{enumerate}

% Question 4
\item[4.] Find the partial total derivatives $\S W / \xi u$ and $\S W / \xi v$ if
\begin{enumerate}
\item $W=a x^{2}+b x y+c u$, where $x=\alpha u+\beta v$ and $y=\gamma u$
\item $W=f\left(x_{1}, x_{2}\right)$, where $x_{1}=5 u^{2}+3 v$ and $x_{2}=u-4 v^{3}$
\end{enumerate}
\end{enumerate}



\end{document}