\documentclass{./../../Latex/homework}
\begin{document}
\thispagestyle{plain}
\newcommand{\mytitle}{Homework 5 Problems}
\myheader{\mytitle}

%%%%%%%%%%%%%% Exercise 6.2
\subsection*{Exercise 6.2}

\begin{enumerate}

% Question 2
\item[2.] Given the function $y=5 x^{2}-4 x$ :
\begin{enumerate}
\item Find the difference quotient as a function of $x$ and $\Delta x$.
\item Find the derivative $d y / d x$. (Using the limit definition.)
\item Find $f^{\prime}(2)$ and $f^{\prime}(3)$.
\end{enumerate}

% Question 3
\item[3.] Given the function $y=5 x-2$ :
\begin{enumerate}
\item Find the difference quotient $\Delta y / \Delta x$. What type of function is it?
\item Since the expression $\Delta x$ does not appear in the function $\Delta y / \Delta x$ in part (a), does it make any difference to the value of $\Delta y / \Delta x$ whether $\Delta x$ is large or small? Consequently, what is the limit of the difference quotient as $\Delta x$ approaches zero?
\end{enumerate}
\end{enumerate}

%%%%%%%%%%%%%% Exercise 7.1
\subsection*{Exercise 7.1}

\begin{enumerate}

% Question 3
\item[3.] Find $f^{\prime}(1)$ and $f^{\prime}(2)$ for the following functions:
\begin{tasks}(3)
\task $y=f(x)=18 x$
\task $y=f(x)=c x^{3}$
\task $f(x)=-5 x^{-2}$
\task $f(x)=\frac{3}{4} x^{4 / 3}$
\task $f(w)=6 w^{1 / 3}$
\task $f(w)=-3 w^{-1 / 6}$
\end{tasks}
\end{enumerate}

%%%%%%%%%%%%%% Exercise 7.2
\subsection*{Exercise 7.2}

\begin{enumerate}

% Question 3
\item[3.] Differentiate the following by using the product rule:
\begin{tasks}
\task[(d)]$(a x-b)\left(c x^{2}\right)$
\task[(e)]	$(2-3 x)(1+x)(x+2)$
\end{tasks}

% Question 7
\item[7.] Find the derivatives of:
\begin{tasks}(2)
\task  $\left(x^{2}+3\right) / x$
\task  $(x+9) / x$
\task  $6 x /(x+5)$
\task  $\left(a x^{2}+b\right) /(c x+d)$
\end{tasks}

% Question 8
\item[8.] Given the function $f(x)=a x+b$, find the derivatives of:
\begin{tasks}(4)
\task $f(x)$
\task $x f(x)$
\task $1 / f(x)$
\task $f(x) / x$
\end{tasks}

\end{enumerate}


%%%%%%%%%%%%%% Exercise 7.3
\subsection*{Exercise 7.3}

\begin{enumerate}

% Question 1
\item Given $y=u^{3}+2 u$, where $u=5-x^{2}$, find $d y / d x$ by the chain rule.

% Question 2
\item Given $w=a y^{2}$ and $y=b x^{2}+c x$, find $d w / d x$ by the chain rule.

% Question 3
\item Use the chain rule to find $d y / d x$ for the following:
\begin{tasks}(3)
\task $y=\left(3 x^{2}-13\right)^{3}$
\task $y=\left(7 x^{3}-5\right)^{9}$
\task $y=(a x+b)^{5}$
\end{tasks}

% Question 4
\item Given $y=(16 x+3)^{-2}$, use the chain rule to find $d y / d x$. Then rewrite the function as $y=1 /(16 x+3)^{2}$ and find $d y / d x$ by the quotient rule. Are the answers identical?

% Question 5
\item Given $y=7 x+21$, find its inverse function. Then find $d y / d x$ and $d x / d y$, and verify the inverse-function rule. Also verify that the graphs of the two functions bear a mirrorimage relationship to each other.

% Question 6
\item Are the following functions strictly monotonic?
\begin{tasks}(2)
\task $y=-x^{6}+5 \quad(x>0)$
\task $y=4 x^{5}+x^{3}+3 x$
\end{tasks}
For each strictly monotonic function, find $d x / d y$ by the inverse-function rule.
\end{enumerate}


\end{document}